### 2020-11-10 latex/sections/background.tex
iteration

In this chapter we discuss traditional artificial intelligence in the context of search the algorithms we used, what some other similar 
methods exist as well as the field in general.

@@

When doing research on Artificial Intelligence it is important to select 
an environment that is a suitable abstraction for the task at hand. Environments vary 
significantly, and can be identified by their characteristic. The characteristics 
that are generally used to describe environments can be seen in Table \ref{tab:env_characteristics}. 
CITE AIMABOOK

\begin{table}[h]
  \centering
  \begin{tabular}{|c|c|p{6cm}|}
    \hline
    \textbf{characteristic} & \textbf{Values} & \textbf{Description}\\
    \hline
    Observable & Fully, Partially & How much of the environment can your agent percieve.\\
    \hline
    Agents & Single, Multi & Are there multiple agents playing in the environment.\\
    \hline
    Deterministic & Deterministic, Stochastic & Do the actions your agent do deterministicly impact the environment\\
    \hline
    Episodic & Sequential, Episodic & Are actions episodic or sequential\\
    \hline
    Static & Static, Semi-Static, Dynamic & ehh  \\
    \hline
    Discrete & Discrete, Continuous & Is your environment discrete w.r.t actions (does it end) \\
    \hline
  \end{tabular}
  \caption{Characteristics of environments}
  \label{tab:env_characteristics}
\end{table}

Categorizing environments like this gives you the power to find an environment in which 
a method works and know it can be applied to different environments with the same characteristics.
Talk about \textbf{agents} in the context of environments as the entities that act within the 
environment. A game player is an agent as well as the neural network model that exists within 
a car that is able to drive in the real world.

@@

The game breakthrough is a simplified version of chess, the game is set up on an $MxN$ board with 
cells like in chess, and each player starts with two rows of pawns at either end. The objective of the 
game is to move one of the pawn pieces to the opposite end of the board. A player wins if 
either they have reached the opposite end of the board, or has captured all of his opponents pawns.
The pawns differ from chess pawns in such a way that they can not move two squares on the first move and,
they can move diagonally as well as forward, this leads to the game being impossible to draw as 
pieces are always able to move. An example of an initial board in breakthrough can be seen in Figure \ref{fig:initbtboard}

@@

\begin{table}[h]
  \centering
  \begin{tabular}{|c|c|}\hline
    \textbf{Characteristic} & \textbf{Value}\\\hline
    Observable & Fully \\
    Agents & Multi \\
    Deterministic & Deterministic\\
    Episodic & Sequential \\
    Static & Static \\
    Discrete & Discrete \\\hline
  \end{tabular}
  \caption{Categorization of Breakthrough}
  \label{tab:breakthrough_cat}
\end{table}

@@

@@

To evaluate the game of Breakthrough we can consider many heuristics (higher-level concepts) for instance 
a very simple heuristic would be the amount of pieces each player has left. This heuristics gives us some 
insight into how well the game is progressing, but obviously there are cases where this doesn't tell us 
much, as in cases when your opponent has a single piece left that is on the row immediately before the 
row needed for him to win. No matter how many pieces you have left, this state is bad for you if you're not 
able to capture that piece. An example of such a state can be seen in Figure \ref{fig:bt_h1_bad}.

@@

A different heuristic would be the distance of your most advanced pawn minus your opponents most 
advanced pawn, this heuristic could give you insight into how close you are to winning the game or 
how close your opponent is. As a higher level concept we can call this concept your aggressiveness, 
as it closely resembles how aggressive you are for going in for the win. Generally in Breakthrough 
it is favorable to move your whole team as a unit and play more defensively (CITE UNITY HEURISTIC), so this heuristic 
is probably not one that will lead to the most wins.

@@

Traditionally methods for playing games were searching through the environment using a heuristic to
guide the search. A simple way of doing a heuristic based search would be to give all states a $0$ 
as a score and give states where you win a positive score. We say that those search algorithm 
is not guided, and the algorithm will probably have to evaluate every state in the state-space. 
This method of searching is generally extremely inefficient as the state-spaces of game environments 
are often extremely large, and sometimes infinite. For instance an estimate of the state-space 
for the game breakthrough is $M * N * 2^{N*4}$ where $M$ is the height of the board $N$ is the width 
of the board. The $2^{N*4}$ represent each piece either being alive or captured. So for a small board
$5x4$ the estimated state-space is $1310720$ many states.

@@

The algorithms that are used in traditional state-space search are for instance Depth-first search (DFS),
Breadth-first Search (BFS), Alpha-Beta Pruning Search (AB-Search), and Monte-Carlo Tree Search (MCTS).

More modernly these search methods have been amplified by Machine Learning, in such a way that 
we do not need to figure out a good heuristic for a given state, but rather, we apply a machine 
learning model to learn a function that takes in a state and returns an evaluation of that state.
This can lead to a significant time reduction as we do not need to simulate a whole game from a 
state to receive its evaluation we rather receive the evaluation from the model.

@@

In the algorithm Monte-Carlo Tree Search there is an agent exists
within some environment. Where each node in the environment represents a state-action pair of the environment, by
state-action pair what is meant it is some state and the action that brought the agent to
that state. This pair should be unique within the environment. We will discuss these concepts within game-environments
and therefore we might say player instead of agent and game instead of environment.

@@

In MCTS there is a tree representing the game-environment. This tree consists of nodes $n_i$ where $i$ represents the point in time
of the node, for example $N_0$ in chess is the initial position and $N_x$ is some position in the middle of the game and $N_e$ is one of
the states representing a position where there is either a draw or one player has won the match. Each of the nodes has $4$ values,
$s$, $a$, $Q$, and $N$. These values represent these items, $s$ is state of the environment, $a$ is the action that brought
the previous node $n_(i-1)$ to node $n_i$, $Q$ is the average value of the node (value meaning the outcome of the game),
and $N$ the amount of times the MCTS algorithm has visited this node. The values $s$ and $a$ uniquely identify a position
in the environment and are often called state-action pairs.

@@

\subsection*{Selection}
During the selection phase a node $(s,a)$ within the tree which has not yet been expanded is found.
This process uses Upper Confidence Bound on Trees (UCT) to find that node $(s,a)$, the formula is described in Equation \ref{UCT_formula}. For a parent node $(s,a)$
(initially the root of the tree) we select the child with the highest UCT value. Repeatedly until an unexpanded
node is found. This process is done to balance the amount of exploration vs exploitation of nodes in the
tree.

@@

\subsection*{Rollout}
Next during rollout actions from $(s,a)$ are randomly selected to move to $(s',a')$, then repeated
to go to $(s'',a'')$, until a terminal node within the environment is reached. By terminal we mean a
state in which the game is finished. A terminal node in MCTS can generally return any value, but in the
context of this paper we only return (+1 white wins, -1 black wins, 0 draw).

@@

When training a neural network the UCT formula is modified slightly to prefer selecting nodes
that the neural network values highly by introducing a second scalar to the formula $f(s,a) = (p,v)$,
the resulting formula is described in Equation \ref{PUCT_formula}, and is called PUCT. Secondly, the backpropagation process
is modified to instead of doing rollout/simulation to recieve a reward the predicted value $v$ from
the neural network is used instead.

@@

Machine Learning (ML) is a research field in which machines apply statistical functions on data to achieve the output that is \textit{correct}, by correct we mean the corresponding result which we expect. Generally this a repetitive process where we look at examples of the data, and the algorithm progressively gets closer to the underlying function of the data it is fed. This process is therefore similar to trial and error for humans. ML is a sub-field of Artificial Intelligence. ML algorithms come in the form of two classes \textit{Classification}, where the algorithm should find a class representing the data its given, and \textit{Regression} where the algorithm should find an underlying continuous numerical function and will return a numerical value representing the input.

@@

Neural networks (NN) are popular methods within a sub-field of ML which is called Deep Learning (DL). 
NN's are created to resemble how the human brain functions. In the brain we have neurons which 
when they get a signal they apply some function to them and if the resulting signal is high 
enough, they fire to the next neuron. This is how it is done in the neural network model as well.
There we have neurons that when they get some input, generally a vector of numbers. The neuron 
takes the sum of that vector, weights the sum by a constant, then applies a non-linear activation 
function to it. The result of doing this is then passed on to the next neuron. Until a final layer 
of neurons is reached. At that point we have a value that the neural network corresponds to the 
input value. This value can be a binary classification (cat or dog image), a regression value (the 
value of a property), or any number of outputs. It can then be said that a neural network is doing 
a function approximation of an input to some value. $f_n ( w_n * f_{n-1} (w_{n-1} * \dots f_0(w_0*i))) = o$.

Neural networks are machine learning models that take in as input anything that is numerical 
and they apply differentiatable function to that input s.t. they end with some output. This 
output is then compared to the expected output the difference between these outputs is called 
a loss. A loss is backpropagated through the neural network, and each function is derived in 
order to find its slope. We can then modify the weights of the neural network in the direction 
of the correct output. Leading to a function approximation of this function
$f_{neuralnetwork}(X_{input}) = O_{expectedoutput}$

@@

The field of XAI research is still very far behind it's counterpart AI research,
within XAI two fields are concidered the largest that is Model Interpretability
and Model Explainability.

@@

\subsubsection{Saliency Maps}
Within XAI many methods have been developed to try to evaluate
ANN's, these methods are often referred as model interpretability methods.
In the field image recognition there has been a lot of work examining which
pixels of an image the model deems important. Arguably the most popular
method for this is Saliency Maps (CITE ME). There the pixel values the model
deems important are colored in s.t. a human can examine the image and get a
sense of what portions of the image are important to the model, an example of
a classification of a dog can be seen in Figure \ref{fig:dog_saliency}.

@@

Methods for explaining models that aren't image recognition models include
shapley values. There the input is examined against it's output, then iteratively
input values are selected to be fixed. Then the other input values are varied and
an average change in prediction is calculated. With this the shapley value can be
estimated for the fixed input value. This is done to examine which input values have
the strongest link to the output value. Shapley values on a dataset can give insights
on which input values the model deems important.

@@

A recent paper by Been Kim Et. al (CITATION NEEDED), shows a method for examining
a neural network giving a much more human insight into a prediction. Using Concept
Activation Vectors (CAV) a directional derivative for a given input can be examined
with respect to some HLC's. For example, when a human looks at an image of an animal
and is supposed to decide whether the image is of a horse or a zebra, an intuitive
approach would be to check whether the animal has stipes, or is both white and black.
That method of determining if a horse is a zebra could then be called a higher-level
concept, and if we're able to gather if a nerual network uses this strategy for prediction 
we have a deeper understanding of its underlying structure. Leading to an explanation of 
the result.

CAV's in neural networks are just binary classifiers of the same dataset used to train
a neural network. The

>>>>

In this chapter, we discuss traditional artificial intelligence in the context of search the algorithms we used, what some other similar methods exist as well as the field in general. We allocate a large portion of the section to Breakthrough as that will be the testbed for future sections

@@

When researching Artificial Intelligence it is important to select an environment that is a suitable abstraction for the task at hand. Environments vary significantly and can be identified by their characteristic. The characteristics that are generally used to describe environments can be seen in Table~\ref{tab:env_characteristics}. \cite{Russell:AIModern}

\begin{table}[h]
  \centering
  \begin{tabular}{|c|c|p{6cm}|}
    \hline
    \textbf{characteristic} & \textbf{Values}              & \textbf{Description}                                                \\
    \hline
    Observable              & Fully, Partially             & How much of the environment can your agent percieve.                \\
    \hline
    Agents                  & Single, Multi                & Are there multiple agents playing in the environment.               \\
    \hline
    Deterministic           & Deterministic, Stochastic    & Do the actions your agent do deterministicly impact the environment \\
    \hline
    Episodic                & Sequential, Episodic         & Are actions episodic or sequential                                  \\
    \hline
    Static                  & Static, Semi-Static, Dynamic & ehh                                                                 \\
    \hline
    Discrete                & Discrete, Continuous         & Is your environment discrete w.r.t actions (does it end)            \\
    \hline
  \end{tabular}
  \caption{Characteristics of environments}
  \label{tab:env_characteristics}
\end{table}

Categorizing environments like this gives you the power to find an environment in which a method works and know it can be applied to different environments with the same characteristics. Talk about agents in the context of environments as the entities that act within the environment. A game player is an agent as well as the neural network model that exists within a car that can drive in the real world.

@@

The game breakthrough is a simplified version of chess, the game is set up on a $MxN$ board with cells like in chess, and each player starts with two rows of pawns at either end. The objective of the game is to move one of the pawn pieces to the opposite end of the board. A player wins if either they have reached the opposite end of the board or have captured all of his opponents' pawns. The pawns differ from chess pawns in such a way that they can not move two squares on the first move and, they can move diagonally as well as forward, this leads to the game being impossible to draw as pieces are always able to move. An example of an initial board in breakthrough can be seen in Figure~\ref{fig:initbtboard}

@@

\begin{table}[h]
  \centering
  \begin{tabular}{|c|c|}\hline
    \textbf{Characteristic} & \textbf{Value} \\\hline
    Observable              & Fully          \\
    Agents                  & Multi          \\
    Deterministic           & Deterministic  \\
    Episodic                & Sequential     \\
    Static                  & Static         \\
    Discrete                & Discrete       \\\hline
  \end{tabular}
  \caption{Categorization of Breakthrough}
  \label{tab:breakthrough_cat}
\end{table}

@@

Categorizing Breakthrough with the characteristics described in Section~\ref{sec:environments}. We end up with the description shown in Table~\ref{tab:breakthrough_cat}. These characteristics are identical to that of Tic-Tac-Toe, and Chess. This categorization is the most common in board games where two-player compete.

@@

To evaluate the game of Breakthrough we can consider many heuristics (higher-level concepts) for instance a very simple heuristic would be the number of pieces each player has left. This heuristics gives us some insight into how well the game is progressing, but obviously, there are cases where this doesn't tell us much, as in cases when your opponent has a single piece left that is on the row immediately before the row needed for him to win. No matter how many pieces you have left, this state is bad for you if you're not able to capture that piece. An example of such a state can be seen in Figure~\ref{fig:bt_h1_bad}.

@@

A different heuristic would be the distance of your most advanced pawn minus your opponents most advanced pawn, this heuristic could give you insight into how close you are to winning the game or how close your opponent is. As a higher-level concept, we can call this concept your aggressiveness, as it closely resembles how aggressive you are going for in for the win. Generally, in Breakthrough, it is favorable to move your whole team as a unit and play more defensively, so this heuristic is probably not one that will lead to the most wins.

@@

Traditionally methods for playing games were searching through the environment using a heuristic to guide the search. A simple way of doing a heuristic-based search would be to give all states a $0$ as a score and give states where you win a positive score. We say that that search algorithm is not guided, and the algorithm will probably have to evaluate every state in the state-space. This method of searching is generally extremely inefficient as the state-spaces of game environments are often extremely large, and sometimes infinite. For instance, an estimate of the state-space for breakthrough is $M * N * 2^{N*4}$ where $M$ is the height of the board $N$ is the width of the board. The $2^{N*4}$ represents each piece either being alive or captured. So for a small board, $5x4$ the upper-bound estimation of the state-space is $1310720$ states.

@@

The algorithms that are used in traditional state-space searches are for instance Depth-first search (DFS), Breadth-first Search (BFS), Alpha-Beta Pruning Search (AB-Search), and Monte-Carlo Tree Search (MCTS).

More modernly, these search methods have been amplified by Machine Learning, in such a way that we do not need to figure out a good heuristic for a given state, but rather, we apply a machine learning model to learn a function that takes in a state and returns an evaluation of that state. This can lead to a significant time reduction as we do not need to simulate a whole game from a state to receive its evaluation we rather receive the evaluation from the model.

@@

In the algorithm Monte-Carlo Tree Search, there is an agent within some environment. Where each node in the environment represents a state-action pair of the environment, by state-action pair what is meant it is some state and the action that brought the agent to that state. This pair should be unique within the environment. We will discuss these concepts within game-environments and therefore we might say player instead of agent and game instead of the environment.

@@

In MCTS there is a tree representing the game-environment. This tree consists of nodes $n_i$ where $i$ represents the point in time of the node, for example, $N_0$ in chess is the initial position and $N_x$ is some position in the middle of the game and $N_e$ is one of the states representing a position where there is either a draw or one player has won the match. Each of the nodes has $4$ values, $s$, $a$, $Q$, and $N$. These values represent these items, $s$ is the state of the environment, $a$ is the action that brought the previous node $n_(i-1)$ to node $n_i$, $Q$ is the average value of the node (value meaning the outcome of the game), and $N$ the number of times the MCTS algorithm has visited this node. The values $s$ and $a$ uniquely identify a position in the environment and are often called state-action pairs.

@@

\subsection*{Selection}
During the selection phase, a node $(s,a)$ within the tree which has not yet been expanded is found.
This process uses Upper Confidence Bound on Trees (UCT) to find that node $(s,a)$, the formula is described in Equation \ref{UCT_formula}. For a parent node $(s,a)$
(initially the root of the tree) we select the child with the highest UCT value. Repeatedly until an unexpanded
node is found. This process is done to balance the amount of exploration vs exploitation of nodes in the
tree.

@@

\subsection*{Rollout}
Next during rollout actions from $(s,a)$ are randomly selected to move to $(s',a')$, then repeated
to go to $(s'',a'')$, until a terminal node within the environment is reached. By terminal, we mean a
state in which the game is finished. A terminal node in MCTS can generally return any value, but in the
context of this paper, we only return (+1 white wins, -1 black wins, 0 draw).

@@

When training a neural network the UCT formula is modified slightly to prefer selecting nodes
that the neural network values highly by introducing a second scalar to the formula $f(s,a) = (p,v)$,
the resulting formula is described in Equation \ref{PUCT_formula}, and is called PUCT. Secondly, the backpropagation process
is modified to instead of doing rollout/simulation to receive a reward the predicted value $v$ from
the neural network is used instead.

@@

Machine Learning (ML) is a research field in which machines apply statistical functions on data to achieve a correct output, by \textit{correct} we mean the corresponding result which we expect. Generally, this a repetitive process where we look at examples of the data, and the algorithm progressively gets closer to the underlying function of the data it is fed. This process is therefore similar to trial and error for humans. ML is a sub-field of Artificial Intelligence. ML algorithms come in the form of two classes \textit{Classification}, where the algorithm should find a class representing the data it is given, and \textit{Regression} where the algorithm should find an underlying continuous numerical function and will return a numerical value representing the input.

@@

Neural networks (NN) are popular methods within a sub-field of ML which is called Deep Learning (DL).
NN's are created to resemble how the human brain functions. In the brain, we have neurons which when they get a signal they apply some function to them and if the resulting signal is high enough, they fire to the next neuron. This is how it is done in the neural network model as well.
There we have neurons that when they get some input, generally a vector of numbers. The neuron takes the sum of that vector, weighs the sum by a constant, then applies a non-linear activation function to it. The result of doing this is then passed on to the next neuron. Until a final layer of neurons is reached. At that point, we have a value that the neural network corresponds to the input value. This value can be a binary classification (cat or dog image), a regression value (the value of a property), or any number of outputs. It can then be said that a neural network is doing a function approximation of the input to some value. $f_n ( w_n * f_{n-1} (w_{n-1} * \dots f_0(w_0*i))) = o$.

Neural networks are machine learning models that take in as input anything numerical and they apply continuous differentiable functions to that input s.t. they end with some output. This output is then compared to the expected output the difference between these outputs is called a loss. A loss is backpropagated through the neural network, and each function is derived to find its slope. We can then modify the weights of the neural network in the direction of the correct output. Leading to a function approximation of this function $f_{neuralnetwork}(X_{input}) = O_{expectedoutput}$

@@

The field of XAI research is still very far behind its counterpart AI research, within XAI two fields are the largest, those are Model Interpretability and Model Explainability.

@@

\subsubsection{Saliency Maps}
Within XAI many methods have been developed to try to evaluate
ANN's, these methods are often referred as model interpretability methods.
In the field image recognition there has been a lot of work examining which
pixels of an image the model deems important. Arguably the most popular
method for this is Saliency Maps\cite{Koch:saliency}. There the pixel values the model
deems important are colored in s.t. a human can examine the image and get a
sense of what portions of the image are important to the model, an example of
a classification of a dog can be seen in Figure \ref{fig:dog_saliency}.

@@

Methods for explaining models that aren't image recognition models include
shapley values, from the Lundberg \& Lee\cite{LundbergL:shapley}. There the input is examined against it's output, then iteratively
input values are selected to be fixed. Then the other input values are varied and
an average change in prediction is calculated. With this the shapley value can be
estimated for the fixed input value. This is done to examine which input values have
the strongest link to the output value. Shapley values on a dataset can give insights
on which input values the model deems important.

@@

A recent paper by Been Kim Et. al \cite{Keem:TCAV}, shows a method for examining
a neural network giving a much more human insight into a prediction. Using Concept
Activation Vectors (CAV) a directional derivative for a given input can be examined
with respect to some HLC's. For example, when a human looks at an image of an animal
and is supposed to decide whether the image is of a horse or a zebra, an intuitive
approach would be to check whether the animal has stipes, or the animal has both white and black colors.
That method of determining if a horse is a zebra could then be called a higher-level
concept, and if we're able to gather if a nerual network uses this strategy for prediction
we have a deeper understanding of its underlying structure. Leading to an explanation of
the result.

The construction of a CAV requires a method of labelling the values in your dataset with
the corresponding concept in order to create a simple binary classifier on the data points.

### 2020-11-10 latex/sections/discussion.tex
iteration

\ifdraft{Here I will discuss the myriad of fields research like this can help
for instance for health organizations, patients, taxpayers, and pharma}
hey

@@

\ifdraft{summarize the work}
hey

@@

\ifdraft{conclude the work, discuss the significance of the work}
hey

>>>>

\ifdraft{Here I will discuss the myriad of fields research like this can help
for instance for health organizations, patients, taxpayers, and pharma}

This section discusses the results and the future work that could span from this research. Additionally, we conclude the work.

@@

As seen in Chapter~\ref{cha:results}. the our trained agent does indeed quickly rise to using the higher level concepts. However, there is not a lot of difference between each heuristic, and not much variability in the heuristics once it reaches $\tilde50\%$. These results are not what we hoped we would see, as a concept like numbers advantage clearly isn't a great heuristic. A reasonable assumption would be that as we train the agent more it will start to use these strategies more, or to increase the size of the neural network to gain performance equal or better to that of a human player.

There is a possibility that the described HLC's are not the HLC's that a neural network uses to play the game, and thus we don't see a connection between the network and the concepts. The amount of HLC however is infinite and they must be selected carefully, we selected only ones that are common beginner heuristics in board games. A possible avenue of research would be to examine a fully trained neural network that does compete in games at a super human level, a network like AlphaZeroGo and examine it's play style. From its play style it would be possible to gather which HLC's it focusess on, then retraining the model examining these HLC's over time to find new pedagogical method of teaching Go.

Additionally applying this method to a greater set of environment would be an interesting field of research. We envision a self driving car agent being examined with respect to a higher level concept of aggressiveness, or a mortgage agent being examined with respect to a higher level concept of racism.

@@

\ifdraft{conclude the work, discuss the significance of the work}

The work done in this paper is only a first step in examining working agents in active environments with respect to human level concepts. If improved could have a great impact in our trust on neural network that improve our daily lives. While the testbed for the research was a discrete game environment, there is a clear way forward to a dynamic system with discrete human level concepts. Our results show that for an agent that clearly doesn't achieve human level performance it's actions are often selected with respect to human level concepts, up to 60\% of the actions.

### 2020-11-10 latex/sections/introduction.tex
iteration

The world of deep learning (DL) is exciting, we have models that are able to examine
images and very reliably be able to identify it's content. Deep learning models have
beaten Opthalmologists in identifing diabetic retinopathy, they've identified
cancer cells where others have not. Deep learning models have even predicted
the likelyhood a person will die in the following year with good accuracy (CITE).
These models have done these things extremely accurately, cheap, and fast.

DL in conjunction with a randomized State-space Search Algorithm Monte-Carlo Tree Search, 
was used to defeat the standing world champion Lee Sedol in the incredibly expansive game of Go. This was done in the year 
2017, the researchers used reinforcement learning (RL), where the agent learned by only playing against itself. 
These result imply that given enough time to learn the computer agent is able to
gain a deep insight on how the game should be played.

The field of examining deep learning models to gain a richer understanding
of how they work, and what makes them so much better than humans at various tasks
is still new. This is the field of Explainable Artificial Intelligence (XAI).
If we wish to continue using artificial intelligence (AI) and machine learning (ML),
to improve our lives we must examine how they work, this is not only to improve
the models themselves but could also augment our ability in many cases.
Furthermore, these explanations are a legal requirement now in many areas, namely, legal, and
medical. Recently, XAI has generally been focused on Model Interpretability, in
particular focus on gaining insights on the concepts learned by Deep Neural Networks.

In this we thesis takes a look at how we can examine an artificial neural network (ANN)
to understand which higher level concepts (HLC) it deems important for a given state
within games. These games can be simple like Tic-Tac-Toe or Breakthrough and extremely
complicated like Chess or Go.

The method of examining HLC's within games has generally been done by examining the
current state of the game by evaluating them with respect to some heuristic. A heuristic within
games are evaluations of the end reward for a state given only the state, for example,
the amount of pieces left within a game of chess. Intuitively, the amount of pieces
left is a good estimate for a state in chess, this is a higher level concept we
use to evaluate a chess position. The piece amount can be considered as a lower-level concept
than other concepts we use. For instance, grand master chess players evaluate a position w.r.t. states
where the king is safe from attack, or the structure of the pawn positions.
This paper examines a neural networks evaluation of a state regarding
those higher-level concepts, if human players evaluate the king safety of a state
low but the neural network highly values it, and the neural network plays better
than the player. There could be an avenue for us to improve our game by considering king safety more thoroughly.

@@

In this thesis we take an example game of Breakthrough, train a neural network to play the game
using only self-play. That is, the neural network is trained only by playing against
itself with no outside input other than the rules of the game itself. And we examine
the neural networks against popular Higher-level concepts (heuristics) that we use
to play the game.

>>>>

The world of deep learning (DL) is exciting, we have models that can examine
images and very reliably be able to identify it's content. Deep learning models have
beaten Ophthalmologists in identifying diabetic retinopathy, they've identified
cancer cells where others have not. Deep learning models have even predicted
the likelihood a person will die in the following year with good accuracy (CITE).
These models have done these things extremely accurately, cheap, and fast.

DL in conjunction with a randomized State-space Search Algorithm Monte-Carlo Tree Search was used to defeat the standing world champion Lee Sedol (CITE ALPHAGO ZERO) in the incredibly expansive game of Go. This was done in the year 2017, the researchers used reinforcement learning (RL), where the agent learned by only playing against itself. 
These results imply that given enough time to learn the computer agent can gain a deep insight into how the game should be played.

The field of examining deep learning models to gain a richer understanding
of how they work, and what makes them so much better than humans at various tasks
is still new. This is the field of Explainable Artificial Intelligence (XAI).
If we wish to continue using artificial intelligence (AI) and machine learning (ML),
to improve our lives we must examine how they work, this is not only to improve
the models themselves but could also augment our ability in many cases.
Furthermore, these explanations are a legal requirement now in many areas, namely, legal, and
medical. Recently, XAI has generally been focused on Model Interpretability (MI), in
particular, interpreting the model in such a way that we might predict what the model will
do given a particular input. Another class of XAI would be Model Explainability (ME), where we can adequately explain how the internal mechanics of the model works.
A simple example of when we're able to apply ME would be when we print out a decision tree
that is generated from an ML algorithm, with the decision tree nodes and is conditions we can see exactly what the tree will do with a given input. While in MI we would have to predict the results, without going into details as to why.

In this thesis, we take a look at how we can examine an artificial neural network (ANN)
to understand which higher-level concepts (HLC) it deems important for a given state
within games. These games can be simple like Tic-Tac-Toe or Breakthrough and extremely
complicated like Chess or Go.

The method of examining HLC's within games has generally been done by examining the
current state of the game by evaluating them concerning some heuristic. A heuristic within games are evaluations of the end reward for a state given only the state, for example,
the number of pieces left within a game of chess. Intuitively, the amount of pieces
left is a good estimate for a state in chess, this is a higher-level concept we
use to evaluate a chess position. The piece amount can be considered as a lower-level concept
than other concepts we use. For instance, grandmaster chess players evaluate a position w.r.t. states
where the king is safe from attack or the structure of the pawn positions.
This paper examines the evaluation of a neural network of a state regarding those higher-level concepts if human players evaluate the king's safety of a state
low but the neural network highly values it, and the neural network plays better
than the player. There could be an avenue for us to improve our game by considering king safety more thoroughly.

@@

In this thesis, we take an example game of Breakthrough, train a neural network to play the game using only self-play. That is, the neural network is trained only by playing against itself with no outside input other than the rules of the game itself. And we examine the neural networks against popular Higher-level concepts (heuristics) that we use to play the game.

### 2020-11-10 latex/sections/managerial.tex
latexing
iteration

%% Dedication is optional, comment out if it is absent
\begin{dedications}
  I dedicate this thesis to my family who while not understanding most of what I do
  always support me. The friends that still want to talk to me even though my
  schedule has rendered me unmeetupwithable. Lastly but certainly not least,
  Hulda Lilja who always easead my mind during my spurts of imposter syndrome, and fear of
  failure.

@@

%% The list of abbreviations is an example of a special list
%% Other lists may be added, such as lists of algorithms, symbols, theorems, etc.
%% IN CS PhD, this is sometimes centered.
\coverchapter{List of Abbreviations}%%RUM: Not mentioned
\begin{tabular}{ll}
MSc &Masters of Science\\
ML & Machine Learning\\
AI & Artificial Intelligence\\
\end{tabular}

>>>>

%% Dedication is optional, comment out if it is absent
\begin{dedications}
  I dedicate this thesis to my family who while not understanding most of what I do
  always support me. The friends that still want to talk to me even though my
  schedule has rendered me unmeetupwithable. Lastly but certainly not least,

@@

%% The list of abbreviations is an example of a special list
%% Other lists may be added, such as lists of algorithms, symbols, theorems, etc.
%% IN CS PhD, this is sometimes centered.
\coverchapter{List of Abbreviations}%%RUM: Not mentioned
\begin{tabular}{ll}
MSc &Masters of Science\\
ML & Machine Learning\\
AI & Artificial Intelligence\\
ANN & Artificial Neural Network\\
DNN & Deep Neural Network\\
MCTS & Monte-Carlo Tree Search\\
CAV & Concept Activation Vector\\
DL & Deep Learning\\
MI & Model Interpretability\\
ME & Model Explainability\\

\end{tabular}
